# Introduction to Data Science HW 4
##### Copyright Jeffrey Stanton, Jeffrey Saltz, and Jasmina Tacheva


```{r}
# Enter your name here: 
```

### Attribution statement: (choose only one and delete the rest)


```{r}
# 1. I did this homework by myself, with help from the book and the professor.
# 2. I did this homework with help from the book and the professor and these Internet sources:
# 3. I did this homework with help from <Name of another student> but did not cut and paste any code.
```

Reminders of things to practice from previous weeks: <br>
Descriptive statistics: 	mean( ) max( ) min( )<br>
Coerce to numeric:		as.numeric( )<br>

## Part 1: Use the Starter Code 

Below, I have provided a starter file to help you. <br>

Each of these lines of code **must be commented** (the comment must that explains what is going on, so that I know you understand the code and results).


```{r}
library(jsonlite)
dataset <- url("https://intro-datascience.s3.us-east-2.amazonaws.com/role.json")
readlines <- jsonlite::fromJSON(dataset)
df <- readlines$objects$person
```

A.	Explore the **df** dataframe (e.g., using head() or whatever you think is best).


```{r}


```

B.	Explain the dataset <br>
o	What is the dataset about? <br>
o	How many rows are there and what does a row represent?<br>
o	How many columns and what does each column represent?<br>



```{r}

```

C. What does running this line of code do? Explain in a comment:


```{r}
vals <- substr(df$birthday,1,4)
```

D. Create a new attribute 'age' - how old the person is 
**Hint:** You may need to convert it to numeric first.


```{r}

```

E. Create a function that reads in the role json dataset, and adds the age attribute to the dataframe, and returns that dataframe


```{r}

```

F. Use (call, invoke) the function, and store the results in df


```{r}

```

## Part 2: Investigate the resulting dataframe 'df'

A.	How many senators are women? 


```{r}

```

B.	How many senators have a YouTube account?  


```{r}

```

C.	 How many women senators have a YouTube account?


```{r}

```

D.	Create a new dataframe called **youtubeWomen** that only includes women senators who have a YouTube account.   


```{r}

```

E.	Make a histogram of the **age** of senators in **youtubeWomen**, and then another for the senetors in **df**. Add a comment describing the shape of the distributions.


```{r}

```
