---
output:
  word_document: default
  html_document: default
---
## Intro to Data Science - HW 1
##### Copyright Jeffrey Stanton, Jeffrey Saltz, and Jasmina Tacheva


```{r}
# Enter your name here: Mark Cappiello
```

### Attribution statement: (choose only one and delete the rest)


```{r}
# 1. I did this homework by myself, with help from the book and the professor.

```

Define a variable:


```{r}
x <- 280

```

Define the following vectors, which represent the **population** (in thousands) and **number of colleges** in each of the five counties in Central New York (CNY) – **Cayuga**, **Cortland**, **Madison**, **Onondaga**, and **Oswego**, in this order:


```{r}
population <- c(80, 49, 73, 467, 122)
colleges <- c(2, 2, 3, 9, 2)
```

# Part 1: Calculating statistics using R  

A.	Show the number of observations in the **population** vector with the length() function: 


```{r}
pop_length <- length(population)
pop_length
```

B.	Show the number of observations in the **colleges** vector with the length() function:


```{r}
col_length <- length(colleges)
col_length
```

C.	Calculate the average CNY population using the mean() function:


```{r}
pop_mean <- mean(population)
pop_mean
```

D.	Calculate the average number of colleges in CNY using the mean() function:


```{r}
col_mean <- mean(colleges)
col_mean
```

E.	Calculate the total CNY population using the sum() function:


```{r}
pop_sum <- sum(population)
pop_sum
```

F.	Calculate the total number of colleges in CNY using the sum() function:


```{r}
col_sum <- sum(colleges)
col_sum
```

G.	Calculate the average CNY population again, this time using **the results from steps A & E**:


```{r}
pop_sum/pop_length
```

H.	Calculate the average number of colleges in CNY again, this time using **the results from steps B & F**:


```{r}
col_sum/col_length
```

# Part 2: Using the max/min and range functions in {r}

I.	How many colleges does the county with most colleges have? Hint: Use the max() function:


```{r}
col_max <- max(colleges)
col_max
```

J.	What is the population of the least populous county in CNY? **Hint:** Use the min() function:


```{r}
pop_min <- min(population)
pop_min
```

K.	Display the populations of the least populous and most populous county in the dataset together. **Hint:** Use the range() function:


```{r}
pop_range <- range(population)
pop_range
```

# Part 3: Vector Math

L.	Create a new vector called **extraPop**, which is the current population of a county **+ 50** (each county has 50,000 more people):


```{r}
extraPop <- population+50
extraPop
```

M.	 Calculate the average of **extraPop**:


```{r}
xtra_mean <- mean(extraPop)
xtra_mean
```

N.	In a variable called **bigCounties**, store all the population numbers from the original **population** vector which are **greater than 120** (using **subsetting** in R):


```{r}
bigCounties <- population > 120
population[bigCounties]
```

O.	Report the length of **bigCounties**:


```{r}
bC_length <- length(population[bigCounties])
bC_length
```
