# Introduction to Data Science HW 4
##### Copyright Jeffrey Stanton, Jeffrey Saltz, and Jasmina Tacheva

```{r}
# Enter your name here: Mark Cappielllo
```

### Attribution statement: (choose only one and delete the rest)

```{r}
# 1. I did this homework by myself, with help from the book and the professor.
```

Reminders of things to practice from previous weeks: <br>
Descriptive statistics: 	mean( ) max( ) min( )<br>
Coerce to numeric:		as.numeric( )<br>

## Part 1: Use the Starter Code 

Below, I have provided a starter file to help you. <br>

Each of these lines of code **must be commented** (the comment must that explains what is going on, so that I know you understand the code and results).


```{r}
#--- the library() function is used to load the jsonlite package so that data in json format can be utilized.
library(jsonlite)
#--- the url() function pulls data in from the web and loads it into a variable called dataset
dataset <- url("https://intro-datascience.s3.us-east-2.amazonaws.com/role.json")
#--- the fromJSON() function from the jsonlite package is used to load data into a variable called readlines
readlines <- jsonlite::fromJSON(dataset)
#--- data from the readlines variable object is loaded into a variable called df
#--- readlines contains data that has nested layers that you see using the str() function
#--- you can also use the names() function so that you can see each layer's column names more clearly
#--- first, names(readlines) this returns "meta" and "objects"
#--- then use names(readlines$objects) to see the columns that are available on that level
df <- readlines$objects$person
```

A.	Explore the **df** dataframe (e.g., using head() or whatever you think is best).

```{r}
#--- the readlines() function helps you to view different levels of this nested dataset.  By executing the following three commands you can see what column exist on each level
names(readlines)
names(readlines$objects)
names(readlines$objects$person)

#--- the head() function lets you see the first five rows of a dataframe to help yo understand what kind of data is inside 
head(df, 5)

#--- the structure() functions tells you more about the data from
str(df)

#--- the summary() function gives you additional info about a dataframe.such as column names, the mode of the data in each column, if the column contains numeric data it will give you some basic statistics about the column as well.
summary(df)

nrow(df)
ncol(df)

```

B.	Explain the dataset <br>
o	What is the dataset about? <br>
o	How many rows are there and what does a row represent?<br>
o	How many columns and what does each column represent?<br>



```{r}
#--- the df dataset contains information about US senators 
#--- there are 100 rows in this data set.  Each row contains information about about a US senator
#--- there are 17 columns and each column contains a list of specific data points about US senators 
```

C. What does running this line of code do? Explain in a comment:


```{r}
#--- the substr() function is used on the birthday column of the df data frame.
#--- it extracts a substring of the value in each row, which in this case is each senator's birthdate
#--- by passing the values 1 and 4 as the second and third arguments, the first four characters will be extracted 
#--- given the format of the date in the column, the first four characters are the year of birth.
vals <- substr(df$birthday,1,4)
```

D. Create a new attribute 'age' - how old the person is 
**Hint:** You may need to convert it to numeric first.


```{r}
current_year <- as.numeric(substr(Sys.Date(),1,4))
vals <- as.numeric(vals)
df$age <- current_year - vals
```

E. Create a function that reads in the role json dataset, and adds the age attribute to the dataframe, and returns that dataframe


```{r}
myDataset <- url("https://intro-datascience.s3.us-east-2.amazonaws.com/role.json")

dfAges <- function(myDataset){
  myReadlines <- jsonlite::fromJSON(myDataset)
  myData <- myReadlines$objects$person
  myYear <- as.numeric(substr(Sys.Date(),1,4))
  myDate<- as.numeric(substr(myData$birthday,1,4))
  myData$age <- myYear - myDate
  return(myData)
}


```

F. Use (call, invoke) the function, and store the results in df


```{r}
df <- dfAges(myDataset)
```

## Part 2: Investigate the resulting dataframe 'df'

A.	How many senators are women? 


```{r}
women <- nrow(df[df$gender=='female',])
women

```

B.	How many senators have a YouTube account?  


```{r}
youtube <- nrow(df[!is.na(df$youtubeid),])
youtube
```

C.	 How many women senators have a YouTube account?


```{r}
women_with_youtube <- nrow(df[df$gender == 'female' & !is.na(df$youtubeid),])
women_with_youtube
```

D.	Create a new dataframe called **youtubeWomen** that only includes women senators who have a YouTube account.   


```{r}
youtubeWomen <- df[df$gender == 'female' & !is.na(df$youtubeid),]
youtubeWomen
```

E.	Make a histogram of the **age** of senators in **youtubeWomen**, and then another for the senetors in **df**. Add a comment describing the shape of the distributions.


```{r}
#--- for this dataset it seems to be skewed so that there are more data points to the left and fewer to the right. There is also a large gap a in one of the intervals on the right where there were no observations. There may be a few outliers in the data. 
hist(youtubeWomen$age)

#for this dataset the histogram has a more symmetrical shape. and a more normal distributions.
hist(df$age)
```
