---
output:
  word_document: default
  html_document: default
---
# Intro to Data Science HW 7
##### Copyright Jeffrey Stanton, Jeffrey Saltz, and Jasmina Tacheva


```{r}
# Enter your name here: Mark Cappiello
```

### Attribution statement: (choose only one and delete the rest)


```{r}
# 2. I did this homework with help from the book and the professor and these Internet sources: I also recently bought a book called "Learning R" written by Richard Cotton published by O'Reilly Media Inc. 2013
```

The chapter on **linear models** (“Lining Up Our Models”) introduces **linear predictive modeling** using the tool known as **multiple regression**. The term “multiple regression” has an odd history, dating back to an early scientific observation of a phenomenon called **“regression to the mean.”** These days, multiple regression is just an interesting name for using **linear modeling** to assess the **connection between one or more predictor variables and an outcome variable**. 


<br>In this exercise, you will **predict Ozone air levels from three predictors**.

A.	We will be using the **airquality** data set available in R. Copy it into a dataframe called **air** and use the appropriate functions to **summarize the data**. 


```{r}
air <- airquality
str(air)
summary(air)
#glimpse(air)
head(air, 5)
```

B.	In the analysis that follows, **Ozone** will be considered as the **outcome variable**, and **Solar.R**, **Wind**, and **Temp** as the **predictors**. Add a comment to briefly explain the outcome and predictor variables in the dataframe using **?airquality**.


```{r}
#?airquality
#--- outcome variable	Ozone	numeric	Ozone (ppb)
#--- predictor	Solar.R	numeric	Solar R (lang)
#--- predictor	Wind	numeric	Wind (mph)
#--- predictor	Temp	numeric	Temperature (degrees F)
#--- predictor	Month	numeric	Month (1--12)
#--- predictor	Day	numeric	Day of month (1--31)


```

C.	Inspect the outcome and predictor variables – are there any missing values? Show the code you used to check for that.


```{r}
na_values <- is.na(air)
head(na_values, 5)
```

D.	Use the **na_interpolation()** function from the **imputeTS package** (remember this was used in a previous HW) to fill in the missing values in each of the 4 columns. Make sure there are no more missing values using the commands from Step C.


```{r}
#install.packages("imputeTS")
library(imputeTS)

air_no_nas <- apply(air, 2, na_interpolation)
air_no_nas <- as.data.frame(air_no_nas)
head(air_no_nas, 5)

na_values2 <- is.na(air_no_nas)
head(na_values2, 5)

```

E.	Create **3 bivariate scatterplots (X-Y) plots** (using ggplot), for each of the predictors with the outcome. **Hint:** In each case, put **Ozone on the Y-axis**, and a **predictor on the X-axis**. Add a comment to each, describing the plot and explaining whether there appears to be a **linear relationship** between the outcome variable and the respective predictor.


```{r}
library(ggplot2)

#--- Scatterplot for Solar.R vs Ozone
ggplot(air_no_nas, aes(x = Solar.R, y = Ozone)) +
  geom_point() + labs(title = "Solar Radiation vs Ozone", 
       x = "Solar Radiation (Langleys)", y = "Ozone (ppb)") +
  geom_smooth(method = "lm", se = FALSE) +
  annotate("text", x = max(air_no_nas$Solar.R, na.rm = TRUE), 
           y = max(air_no_nas$Ozone,na.rm = TRUE),
           label = "No linear relationship", hjust = 1, vjust = 1)

# Scatterplot for Wind vs Ozone
ggplot(air_no_nas, aes(x = Wind, y = Ozone)) +
  geom_point() + labs(title = "Wind Speed vs Ozone",
       x = "Wind Speed (mph)", y = "Ozone (ppb)") +
  geom_smooth(method = "lm", se = FALSE) +
  annotate("text", x = min(air_no_nas$Wind, na.rm = TRUE), 
           y = max(air_no_nas$Ozone, na.rm = TRUE),
           label = "Possible negative linear relationship", hjust = 0, vjust = 1)

# Scatterplot for Temp vs Ozone
ggplot(air_no_nas, aes(x = Temp, y = Ozone)) +
  geom_point() + labs(title = "Temperature vs Ozone",
       x = "Temperature (Fahrenheit)", y = "Ozone (ppb)") +
  geom_smooth(method = "lm", se = FALSE) +
  annotate("text", x = max(air_no_nas$Temp, na.rm = TRUE), 
           y = max(air_no_nas$Ozone, na.rm = TRUE),
           label = "Possible positive linear relationship", hjust = 1, vjust = 1)

```

F.	Next, create a **simple regression model** predicting **Ozone based on Wind**, using the **lm( )** command. In a comment, report the **coefficient** (aka **slope** or **beta weight**) of **Wind** in the regression output and, **if it is statistically significant**, **interpret it** with respect to **Ozone**. Report the **adjusted R-squared** of the model and try to explain what it means. 


```{r}
model <- lm(Ozone ~ Wind, data = air_no_nas)
summary(model)

# Comment reporting the coefficient of Wind and its interpretation
# The coefficient of Wind is the slope of the regression line, which represents the change in Ozone for a one-unit increase in Wind.
# If the coefficient is statistically significant (p-value < 0.05), we can interpret it as follows:
# For example, if the coefficient of Wind is -2.34 and statistically significant, it means that for every one-unit increase in Wind speed, Ozone decreases by 2.34 parts per billion (ppb) on average.

# Report the adjusted R-squared of the model and its interpretation
# The adjusted R-squared is a measure of how well the independent variable(s) explain the variation in the dependent variable, adjusted for the number of predictors in the model.
# It ranges from 0 to 1, with higher values indicating a better fit.
# For example, an adjusted R-squared of 0.45 means that 45% of the variance in Ozone is explained by Wind in the model.
```

G.	Create a **multiple regression model** predicting **Ozone** based on **Solar.R**, **Wind**, and **Temp**.<br> **Make sure to include all three predictors in one model – NOT three different models each with one predictor.**


```{r}
multiple_model <- lm(Ozone ~ Solar.R + Wind + Temp, data = air_no_nas)
summary(multiple_model)

```

H.	Report the **adjusted R-Squared** in a comment – how does it compare to the adjusted R-squared from above? Is this better or worse? Which of the predictors are **statistically significant** in the model? In a comment, report the coefficient of each predictor that is statistically significant. Do not report the coefficients for predictors that are not significant.


```{r}
# Print the adjusted R-squared of the multiple regression model
# The adjusted R-squared measures the proportion of variance in the dependent variable that is explained by the independent variables in the model, adjusted for the number of predictors.
# Compare it to the adjusted R-squared from the simple regression model.
# If it's higher, it indicates that the multiple regression model explains more variance in Ozone than the simple regression model.
# If it's lower, it indicates that the multiple regression model may not provide a better fit than the simple regression model.

# Print the summary of the multiple regression model to identify "statistically significant" predictors
summary(multiple_model)

# Report the coefficients of statistically significant predictors in a comment
# Statistically significant predictors have p-values less than the chosen significance level (e.g., 0.05).

# Report the coefficients of significant predictors and their interpretation.
# For example, if Solar.R has a p-value less than 0.05, you can report its coefficient and interpret it accordingly.
# Ignore coefficients of predictors that are not "statistically significant".


```

I.	Create a one-row data frame like this: 


```{r}
predDF <- data.frame(Solar.R=290, Wind=13, Temp=61)
```

 and use it with the **predict( )** function to predict the **expected value of Ozone**:


```{r}
predicted_Ozone <- predict(multiple_model, newdata = predDF)
predicted_Ozone
```

J.	Create an additional **multiple regression model**, with **Temp** as the **outcome variable**, and the other **3 variables** as the **predictors**. 

Review the quality of the model by commenting on its **adjusted R-Squared**.  


```{r}
temp_model <- lm(Temp ~ Solar.R + Wind + Ozone, data = air_no_nas)
summary(temp_model)

# Comment on the adjusted R-squared of the model
# The adjusted R-squared measures the proportion of variance in the dependent variable (Temp) that is explained by the independent variables (Solar.R, Wind, and Ozone), adjusted for the number of predictors.

# Interpret the value of adjusted R-squared to assess the quality of the model.
# A higher adjusted R-squared indicates that the model explains more variance in Temp, while a lower adjusted R-squared suggests that the model may not provide a good fit to the data.
```
