---
title: "Data Science for Business With R - CH02"
output:
  html_document:
    df_print: paged
---

## CREATING DATA FRAMES 

First create vectors for the data frame

```{r}
myFamilyNames <- c('Dad', 'Mom','Sis','Bro','Dog')
myFamilyAges <- c(43,42,12,8,5)
myFamilyGenders <- c('Male','Female','Female','Male','Female')
myFamilyWeights <- c(188,136,83,61,44)
```

Create the data frame from the above vectors

```{r}
myFamily <- data.frame(myFamilyNames,myFamilyAges,myFamilyGenders,myFamilyWeights)
myFamily
```
## EXPLORING DATA FRAMES

structure function    

```{r}
str(myFamily)
```

summary function

```{r}
summary(myFamily)
```

head and tail functions show limited results

```{r}
head(myFamily, 3)

tail(myFamily, 3)
```

## ACCESSING COLUMNS IN A DATA FRAME

To view on row of the data frame

```{r}
myFamily$myFamilyAges
```

Adding a value to one of the original vectors does not change the data frame

```{r}
myFamilyAges <- c(myFamilyAges, 11)
myFamilyAges

myFamily$myFamilyAges
```

- myFamily\$myFamilyAges <- c(myFamily\$myFamilyAges, 11)
    - if you try to add an element to one row of an existing data frame
    - It will cause an error and the myFamily data frame will remain changed
    - Data frames must remain rectangular

## CASE STUDY: calculating NPS using a data frame (net promoter score)

```{r}
#--- Six surveys
ltr <- c(9,9,7,6,8,7)
TypeOfTravel <- c('Business Travel','Business Travel','Business Travel',' Mileage',
                   'Personal Travel','Personal Travel')
survey <- data.frame(ltr, TypeOfTravel)

#--- look at the newly created data frame
str(survey)

#--- Calculate the number of promoters and detractors
numP <- sum(survey$ltr > 8)
numD <- sum(survey$ltr < 7 )

#--- Now calculate NPS
total <- nrow(survey)
nps <- (numP/total - numD/total) * 100
nps

survey

#--- do the same analysis, but for business travel tickets
busTravelDF <- survey[survey$TypeOfTravel == "Business Travel",]

#--- Calculate number of promoters and detractors
numP <- sum(busTravelDF$ltr > 8)
numD <- sum(busTravelDF$ltr < 7 )

#--- Calculate Business Travel NPS
total <- nrow(busTravelDF)
bus.nps <- (numP/total - numD/total) * 100
bus.nps
```



```{r}
str(survey$TypeOfTravel) #--- expected output is different in the book

levels(survey$TypeOfTravel) #--- expected output is different in the book
```

```{r}
nps

bus.nps
```
## CHAPTER CHALLENGES

1. Use c() to create new variable myFoods containing five values.

```{r}
myFoods <- c('Pizza','Pizza','Steak','Cereal','Steak')

str(myFoods)
```

2. add myFoods to the myFamily data frame

```{r}
myFamily$myFoods <- myFoods
```

3. Run the summary() function on the new myFamily

```{r}
summary(myFamily)
#--- myFamily$myFoods is data type or mode character
```

4. create an expression that shows a list of TRUE and FALSE values. show TRUE when myAge is less than 40 
assign results to myIndex

```{r}
myIndex <- myFamily$myFamilyAges < 40
```

5. use myIndex to show the favorite food of each kid

```{r}
myFamily$myFoods[myIndex]
```

6. invert the selection using ! to show favorite foods of adults

```{r}
myFamily$myFoods[!myIndex]
```

